---
title: "lab_12"
author: "Derek Willis"
date: "2023-05-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## You will need

* tidytext and our usual libraries

## Load libraries and establish settings

**Task** Create a codeblock and load appropriate packages and settings for this lab.

```{r}
library(tidyverse)
library(tidytext)
library(janitor)
library(lubridate)
library(rvest)
```

## Questions

**Q1.** You've been assigned to report a story about the leading reasons that Maryland attorneys get sanctioned by the state for misconduct. The state [publishes lists of sanctions](https://www.courts.state.md.us/attygrievance/sanctions) that contain a short text description about the situation. Load the CSV file in the data folder containing records from fiscal year 2011 onwards. Make a list of unique words from the text column, then following the example in the pre_lab, remove common "stop words" from that list and create a list of the top 10 words containing the percentage of occurrences each word represents. What's the leading word in that answer and, broadly, what do you think the top 10 words describe?

**My top word is "trust" after removing my stop words. The words remaining, including prejudicial, funds, account, trust, and more, seem to describe the most common reasons that attorneys face sanctions.**  

```{r}
##LOADING DATA
md_attorney_sanctions <- clean_names(read_csv("data/md_attorney_sanctions.csv"))

md_attorney_sanctions
```
```{r}
##FINDING UNIQUE WORDS
unique_words <- md_attorney_sanctions |> 
  select(text) |>
  unnest_tokens(word, text)

unique_words |>
   count(word, sort = TRUE) |>
  top_n(50) |>
  mutate(word = reorder(word, n))
  
```
```{r}
##CREATING STOP WORDS 
data("stop_words")
stop_words <- stop_words |>
  add_row(word="failing") |>
  add_row(word="failed") |>
  add_row(word="engaging") |>
  add_row(word="respondent") |>
  add_row(word="reprimand") |>
  add_row(word="disbarred") |>
  add_row(word="suspension") |>
  add_row(word="counsel") |>
  add_row(word="bar") |>
  add_row(word="justice") |>
  add_row(word="court") |> 
  add_row(word = "attorney") |> 
  add_row(word = "client") |> 
  add_row(word="clients") |> 
  add_row(word="representation") 
```

```{r} 
unique_words |>
  anti_join(stop_words) |>
  group_by(word) |>
  tally(sort=TRUE) |>
  mutate(percentage = (n/sum(n))*100) |>
  top_n(10)
```
**Q2.** Let's move beyond single words to phrases. Make a list of the top 10 three-word phrases, called trigrams, based on the example from the pre_lab (you'll need to modify the example code to do this). What's the top trigram and how often does it appear? What does that phrase mean in legal terms?

**My top outcome was "conduct involving dishonesty." According to an internet search, this is a form of professional misconduct that fits into Rule 8.4 in the ABA, which involves conduct.** 

```{r}

##TOOK MOST OF THIS CODE FROM THE PRE LAB, ADJUSTED FOR MAKING IT A TRIGRAM INSTEAD OF A BIGRAM
##HAD TO GIVE THIS A DIFFERENT NAME SO I COULD COME BACK TO THE FISCAL YEAR ELEMENT FOR THE NEXT QUESTION; WHEN I ORIGINALLY DID THIS, I GAVE IT THE SAME NAME AND IT TOOK OUT FISCAL YEAR
md_attorney_sanctions_trigrams <- md_attorney_sanctions |>
  unnest_tokens(trigram, text, token="ngrams", n=3) |>
  separate(trigram, c("word1", "word2", "word3"), sep = " ") |>
  filter(!word1 %in% stop_words$word) |>
  filter(!word2 %in% stop_words$word) |>
  filter(!word3 %in% stop_words$word) |>
  mutate(trigram = paste(word1, word2, word3, sep=" ")) |>
  group_by(trigram) |>
  tally(sort=TRUE) |>
  mutate(percent = (n/sum(n))*100) |>
  top_n(10)
```
```{r}
md_attorney_sanctions_trigrams
```

**Q3.** Let's drop back down to more traditional text analysis - take the top trigram from Q2 and write code to see how many times it occurs in the text column in each fiscal year. What do you think the answer produced by your code suggests? What else could you do to try and clarify the most important reasons attorneys get sanctioned?

**My results don't seem to show a discernable pattern. There is a spike between 2019 and 2022, but I can't think of why (I even asked my dad, who is an attorney). I think to clarify my results, I should probably take a harder look at my stop words to see if I am leaving anything out that might make more sense. ** 

```{r}
top_result <- "conduct involving dishonesty"

conduct_dishonesty_by_year <- md_attorney_sanctions |>
  filter(str_detect(text, top_result)) |>
  group_by(fiscal_year) |>
  summarise(count=n()) |>
  arrange(desc(fiscal_year))

conduct_dishonesty_by_year
```
