---
title: "lab_03"
author: "Derek Willis"
date: "2024-08-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse
library(tidyverse)

# Load janitor
library(janitor)
```

## Load data

```{r}
# Load MD slice of FEC House candidate contribution data
md_contribs <- read_rds("data/lab_03.rds")
```

## Fix header issues

You'll need to fix one issue with the column names using janitor. Write a code block to do that, making sure that you save the results to the same dataframe.

```{r}
md_contribs <- md_contribs |> 
  clean_names()

```


## Answer questions

Q1: Which seven cities in Maryland had the largest *number* of contributions from donors? How many did each have? To answer this question, you'll have to clean the messy "contributor_city" field.

Note: To answer this question, you'll need to see every variation of the "contributor_city" for the top seven cities, which you should identify through some initial analysis. Then find variations of them and change those variations into the most common version.

A1:The seven cities that saw the most contributions from donors were Bethesda, Chevy Chase, Baltimore, Annapolis, Potomac, Silver Spring, Rockville and Ellicott City. They have 1,117, 704, 692, 493, 491, 408, 359 and 174 (oops I did 8).

```{r}
md_contribs |> 
  group_by(contributor_city) |>
  summarise(contributions = n()) |>
  arrange(desc(contributions))
```

```{r}
md_contribs |> 
    mutate(contributor_city = case_when(
    contributor_city == "BALRIMORE" | contributor_city == "BLATIMORE" ~ "BALTIMORE",
      contributor_city == "ANAPOLIS" | contributor_city == "ANNAPOLIS JUNCTION" ~ "ANNAPOLIS",
    contributor_city == "ELLICOTT" ~ "ELLICOTT CITY",
    contributor_city == "POTAMIC" ~ "POTOMAC",
    contributor_city == "ROCKWELL" ~ "ROCKVILLE",
    contributor_city == "SILVER SPRINGS" ~ "SILVER SPRING",
    TRUE ~ contributor_city
    
  )) |>
  group_by(contributor_city) |> 
  summarise(contributions = n()) |>
arrange(desc(contributions))



```

Q2: Find the 5-digit ZIP code in Maryland that had the highest total amount of contributions (by dollar value). You will need to use mutate to ensure that you have a zip code column with the first five characters from the existing zip code column.

Where is that zip code and what kind of area is it, demographically? Which committee is the largest recipient of money from that zip code? How many of the top five recipients are Maryland candidates?

A2: The zip code with the most is 20815. This applies to Chevy Chase, which is one of the richest areas of Maryland.The largest recipient of money from the zip is Elfreth for MD. All of the top 5 are MD candidates.

```{r}
cleaned_md_contribs <- md_contribs |>
  mutate(contributor_zip = str_sub(contributor_zip, start=1L, end=5L))

cleaned_md_contribs
```
```{r} 
cleaned_md_contribs |> 
  group_by(committee_name, contributor_zip = "20815") |>
summarise(contributions = n()) |>
  arrange(desc(contributions))

```
```{r} 
cleaned_md_contribs |> 
  group_by(contributor_zip) |>
summarise(contributions = n()) |>
  arrange(desc(contributions))
```
Q3: Let's look at top individual donors. Write code to generate a list of top donors using name and city, adding up the total amount of contributions. What does the original data say about the top two donors - who are they and what do they do? Do you see anything in the list of top 10 donors that would make you uncertain about the accuracy of these results?

A3:The top two donors are Daniel Kracov and Margaret Allen. They both work in finance. I don't see anything that makes me uncertain about accuracy, except maybe thart Kracov doubles anyone else.

```{r}
cleaned_md_contribs |> 
  group_by(contributor_name, contributor_city, contributor_employer) |>
summarise(contributions = n()) |>
  arrange(desc(contributions))
```
